---
output:
  html_document: default
  pdf_document:
    keep_tex: yes
---

# Generalized linear models {#chapter12}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
library(mosaic)
library(knitr)
library(gridExtra)
knitr::opts_chunk$set(cache = F,fig.width = 10,  fig.height = 6,
                      out.width="75%",fig.retina = 2, fig.align = 'center',
                      fig.pos='ht!',
                      message=F,warning=F)
options(show.signif.stars = FALSE)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
colFmtBold = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{\\textbf{",x,"}}",sep="")
  else if(outputFormat == 'html')
    paste("<b><font color='",color,"'>",x,"</font></b>",sep="")
  else
    x
}
```


## Situation {#section12-1}

In Chapter \@ref(chapter4), different types of responses and models for them were discussed. This chapter digs into details and provides of each model. This is meant to be a companion with R code to accompany Chapters ... in the _Statistical Sleuth_ [cite]. The treatment in this chapter will review important aspects of the methods, but focus more on code, visualizations, and output for methods covered in the _Sleuth_. 

## Binary logistic regression {#section12-2}


## GLM estimation and the woes of logistic regression {#section12-3}

## Grouped binomial regression {#section12-4}


## Poisson and Poisson rate regression {#section12-5}


## Overdispersion and separation in count models {#section12-6}

## Zero-truncated/zero-inflated models {#section12-7}


<!-- \sectionmark{Multiple (pair-wise) comparisons using Tukey's HSD and CLD} -->

## Multiple (pair-wise) comparisons using Tukey's HSD and the compact letter display {#section12-8}

<!-- \sectionmark{Multiple (pair-wise) comparisons using Tukey's HSD and CLD} -->


\indent There is a method that many researchers use to more efficiently generate and 
report these sorts of results that is called a ***compact letter display*** \index{compact letter display}
(CLD, @Piepho2004)^[Note that this method is implemented slightly differently than explained here in some software packages so if you see this in a journal article, read the discussion carefully.]. The ``cld`` function can be applied to the results from 

<!-- \newpage -->

